---
html_document:
  theme: spacelab
  highlight: tango
  toc: yes
  toc_float:
    collapsed: no
    smooth_scrool: yes
  number_sections: yes
author: "Philipp Pecher"
date: "`r Sys.time()`"
abstract: "In this project I estimated the Unconditional Standard Deviation and Value-at-Risk (VaR) of portfolio consisting four financial instruments with Models of the General Autoregressive Conditional Heteroscedacticity (GARCH) Family (Engele, 2001). The portfolio consists of the follwoing three indexes: Standard and Poor's 500 (SP500), the Deutsche Aktion Index (DAX) and the Warszawski Indeks Giełdowy (WIG20), as well as one cryptocurrency, Bitcoin (BTC). The choice of the Models from the GARCH Family was made on the basis of an comparison analysis. After the comparison and further analysis I choosed an AR(1)-GARCH-t(2,1) and GARCH(5,5). The long-term annulaized unconditional standard deviation is 24.08% (AR(1)-GARCH-t(2,1)) and 12.17 (GARCH(5,5)) for in-sample data. For the out-sample data the number of days were losses are higher than assued VaR are 0.004 (AR(1)-GARCH-t(2,1)) and 0.008 for(GARCH(5,5)). " 
output:
  pdf_document: default
  html_document: default
title: 'Time Series Analysis - Home Project # 2'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE, 
                      cache   = TRUE,
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 10)
```

```{r, include=FALSE}
##############################################################
# Philipp Pecher                                             #             
# p.pecher@student.uw.edu.pl                                 #
# Faculty of Economic Sciences, University of Warsaw         #
#                                                            #
# Time Series Analysis, Spring 2021                          #
#                                                            # 
# Time Series Analysis - Home Project # 2                    #
##############################################################
```

# Time Series Analysis - Home Project #2 Results

## Estimation Garch-Models with  in-sample data

### Packages and data
```{r cache = F}
library(tidyverse)
library(xts) # e.g. xts objects
library(fBasics) # e.g. basicStats()
library(tseries)# e.g. jarque.bera.test()
library(car) # e.g. durbinWatsonTest()
library(FinTS) # e.g. ArchTest()
library(fGarch) # e.g. garchFit()
library(quantmod) # e.g. getSymbol()
library(rugarch) # e.g. ugarchfit()
```


### Functions
```{r}
setwd("/Users/philipp/Library/Mobile Documents/com~apple~CloudDocs/Uni/Time Series Analysis/Project")

source("function_compareICs.GARCH.R")
source("function_compare.ICs.ugarchfit.R")
```

### Choice of cryptocurrency

```{r}
x <- c(9,1,2,7,9,7)

sum(x)%%5+1
```
The result is 1, hence the choice of cryptocurrency is Bitcoin (BTC)

### Analysis of Portfolio

```{r}
SP500 <- 
  getSymbols(Symbols = "^GSPC",             
             from = "2018-01-02", 
             to = "2021-06-01",   
             auto.assign = FALSE)

DAX <- 
  getSymbols(Symbols = "^GDAXI",             
             from = "2018-01-02", 
             to = "2021-06-01",   
             auto.assign = FALSE)

WIG20 <- 
  getSymbols(Symbols = "^WIG20",             
             from = "2018-01-02", 
             to = "2021-06-01",   
             auto.assign = FALSE)

BTC <- 
  getSymbols(Symbols = "BTC-USD",             
             from = "2018-01-02", 
             to = "2021-06-01",   
             auto.assign = FALSE)
```


```{r}
SP500 <- SP500[, 6]
names(SP500) <- "SP500"

DAX <- DAX[, 6]
names(DAX) <- "DAX"

BTC <- BTC[, 6]
names(BTC) <- "BTC"

WIG20 <- WIG20[, 6]
names(WIG20) <- "WIG20"
```

```{r}
SP500$SP500.r <- diff.xts(log(SP500$SP500))

DAX$DAX.r <- diff.xts(log(DAX$DAX))

BTC$BTC.r <- diff.xts(log(BTC$BTC))

WIG20$WIG20.r <- diff.xts(log(WIG20$WIG20))

```

The gathered quotes of stock indices and cryptocurrencies are put togehter in a portfolio with equal shares.
```{r}
quotes <- 0.25*SP500$SP500.r+0.25*DAX$DAX.r+0.25*BTC$BTC.r+0.25*WIG20$WIG20.r
names(quotes) <- "quotes.r"
quotes$quotes <- 0.25*SP500$SP500+0.25*DAX$DAX+0.25*BTC$BTC+0.25*WIG20$WIG20

quotes$obs <- 1:length(quotes$quotes)

```

### Plot of quotations and returns.
```{r}


par(mfrow = c(2, 1))
plot(quotes$quotes.r,
     type = "l", col = "red", lwd = 1,
     main = "Portfolio returns")
plot(quotes$quotes,
     type = "l", col = "black", lwd = 1,
     main = "Portfolio quotes")
```

Examining the returns of the portfolio we see changes between 0.05 and -0.05 except in Febuary 2018 there is a return of -0.06 and in March 2020 there are rapid changes in return from approx. 0.07 to approx -0.21. The COVID-19 pandemic and its uncertainty for the financial market could be a reason for the big differences in returns of our portfolio in March 2020.

### Plot of quotations and returns 
```{r}

par(mfrow = c(2, 1))
plot(quotes$quotes.r,
     type = "l", col = "red", lwd = 1,
     main = "Portfolio returns")
plot(quotes$quotes,
     type = "l", col = "black", lwd = 1,
     main = "Portfolio prices")
par(mfrow = c(1, 1))
```

The behavior of our portfolio prices supports this assumption. The price drops in March where the COVID-19 pandemic and restriction occur in Europe and USA. Nevertheless, the prices recover fast and increase even more. The fast and high increase as well as the drop of prices from December can be cause by Bitcoin which has a similar behavior during this time.

### Plot of ACF for returns 
```{r}

acf(quotes$quotes.r, lag.max = 36, na.action = na.pass,
    col = "darkblue", lwd = 7,
    main = "ACF for Portfolio returns")

acf(quotes$quotes.r, lag.max = 36, na.action = na.pass,
    ylim = c(-0.1, 0.1), 
    col = "darkblue", lwd = 7,
    main = "ACF for portfolio returns (rescaled vertical axis)")



```

The plot of the ACF for my portfolio returns shows to some extent an autoregressive relationship. This can be used to build an ARIMA Model

### Plot of ACF for squared returns 
```{r}

acf(quotes$quotes.r^2, lag.max = 36, na.action = na.pass,
    col = "darkblue", lwd = 7,
    main = "ACF for Portfolio squared returns")

acf(quotes$quotes.r^2, lag.max = 36, na.action = na.pass,
    ylim = c(0, 0.5), 
    col = "darkblue", lwd = 7,
    main = "ACF for Portfolio squared returns (rescaled vertical axis)")

```
The plot of the ACF of squares of my portfolio shows a autoregressive relationship between each other. Thus we can use it to build a GARCH Model.


```{r}
basicStats(quotes$quotes.r) 
hist(quotes$quotes.r, prob = T, breaks = 40, main = "Density of Portfolio log-returns", xlab = "Portfolio log-returns")
curve(dnorm(x, mean = mean(quotes$quotes.r, na.rm = T),
            sd  = sd(quotes$quotes.r, na.rm = T)),
      col = "darkblue", lwd = 2, add = TRUE)
```

The comparison of the distribution of our portfolio returns with a normal distribution shows a highly leptokurtic distribution. 


```{r}
basicStats(quotes$quotes.r)

```

The basic Statistics show among others a Skewness of -2.62, which says, that the left tail is thicker than the right tail of the distribution. The Kurtosis which is the excess Kurtosis of 34.81 suggest a thicker thail than the normal distribution. The Kurtosis would be at approx 39.

```{r}
jarque.bera.test(na.omit(quotes$quotes.r))

```

Additionally, the p-value of the Jarque-Bera test is approx 0.00000000000000002. Hence, we can reject the Null-Hypothesis about normality to a confidence level of 99%.

```{r}
ArchTest(quotes$quotes.r, 
         lags = 5)

ArchTest(quotes$quotes.r, 
         lags = 50)
```
The result of the Arch Test has a p-value of 0.000009 for 5 lags and 0.003 for 50 lags. Thus we can reject the Null-Hypothesis for both to a confidence level of 99% which says that the residuals exhibits no conditional heteroscedasticity (Arch Effects). This result suggest the application of an ARCH Model.


```{r}
quotes$quotes.r[is.na(quotes$quotes.r)] <- 0
```
Replace NA quotes with 0, since the ugarfit() function does not accept missing values.


### Estimation GARCH Models

### Garch(5,5)
```{r}
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", 
                        garchOrder = c(5, 5)),
  mean.model = list(armaOrder = c(0, 0), 
                    include.mean = TRUE), 
  distribution.model = "norm")

quotes$quotes.r[is.na(quotes$quotes.r)] <- 0


k.garch55 <- ugarchfit(spec = spec, 
                           data = quotes$quotes.r)

k.garch55
```

### The AR(5)-EGARCH(2,1) model
```{r}
  spec <- ugarchspec(
    variance.model = list(model = "eGARCH", 
                          garchOrder = c(2, 1)),
    mean.model = list(armaOrder = c(5, 0), 
                      include.mean = TRUE), 
    distribution.model = "norm")
  

  k.ar5egarch21 <- ugarchfit(spec = spec, 
                             data = quotes$quotes.r)
  
  k.ar5egarch21
```

### AR(1)-EGARCH(2,1)
```{r}
  
  spec <- ugarchspec(
    variance.model = list(model = "eGARCH", 
                          garchOrder = c(2, 1)),
    mean.model = list(armaOrder = c(1, 0), 
                      include.mean = TRUE), 
    distribution.model = "norm")
  

  
  k.ar1egarch21 <- ugarchfit(spec = spec, 
                             data = quotes$quotes.r)
  
  k.ar1egarch21


```

### AR(1)-EGARCH(2,1)-mean
```{r}

spec <- ugarchspec(
  variance.model = list(model = "sGARCH", 
                        garchOrder = c(2, 1)),
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE,
                    archm = TRUE, archpow = 1), 
  distribution.model = "norm")


k.ar1garchm21 <- ugarchfit(spec = spec, 
                           data = quotes$quotes.r)

k.ar1garchm21

```


### AR(1)-GARCH-t(2,1)
```{r}
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", 
                        garchOrder = c(2, 1)),
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE), 
  distribution.model = "std") 

k.ar1garcht21 <- ugarchfit(spec = spec, 
                           data = quotes$quotes.r)

k.ar1garcht21
```

### AR(1)-EGARCH-t(2,1)-mean
```{r}
spec <- ugarchspec(
  variance.model = list(model = "eGARCH", 
                        garchOrder = c(2, 1)),
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE,
                    archm = TRUE, archpow = 1), 
  distribution.model = "std") 


k.ar1egarchmt21 <- ugarchfit(spec = spec, 
                             data = quotes$quotes.r)

k.ar1egarchmt21
```

### AR(1)-EGARCH-t(2,1)
```{r}
spec <- 
  ugarchspec(variance.model = list(model = "eGARCH", garchOrder = c(2, 1)),
             mean.model = list(armaOrder = c(1, 0), include.mean = F),
             distribution.model = "std")


k.ar1egarcht21 <- ugarchfit(spec = spec, 
                             data = quotes$quotes.r)

k.ar1egarcht21

```


### Comparison

```{r}
compare.ICs.ugarchfit(c("k.garch55",
                        "k.ar5egarch21",
                        "k.ar1egarch21", 
                        "k.ar1garchm21", 
                        "k.ar1garcht21", 
                        "k.ar1egarchmt21",
                        "k.ar1egarcht21"))

```

The comparison considers four different criteria (Akaike, Bayes, Shibata and Hannan.Quinn). Two of the criteria (Akaike and Shibata) suggest an AR(1)-EGARCH-t(2,1)-mean model and the other two (Bayes and Hannan.Quinn) suggest a AR(1)-GARCH-t(2,1) model. 

### Results
(a) estimates of annualized conditional standard deviation in the in-sample period produced by the two models

### Annualized conditional standard deviation AR(1)-GARCH-t(2,1):
```{r}
sigma.forecast.longrun1 <- ugarchforecast(k.ar1garcht21, n.ahead = 252)
unconditional_sigma1 <-
  sqrt(
    k.ar1garcht21@model$pars["omega", 1] /
      (1 -
         k.ar1garcht21@model$pars["alpha1", 1] -
         k.ar1garcht21@model$pars["beta1", 1]))

plot(
  c(as.numeric(k.ar1garcht21@fit$sigma * sqrt(252)),
    as.numeric(sigma.forecast.longrun1@forecast$sigmaFor * sqrt(252))),
  main = "Annaulized Conditional Standard Deviation GARCH t-model",
  type = "l",
  ylab = "sigma annualized")
abline(h = unconditional_sigma1 * sqrt(252), col = "red")




max(k.ar1garcht21@fit$sigma) * sqrt(252)*100

unconditional_sigma1 * sqrt(252)*100



```
In the plot we see the annualized conditional standard deviation of AR(1)-GARCH-t(2,1). The conditional sigma should converge to to the long run estimate of sigma, which is not the case. This is in contrast to the conventional literature. The deviation could have many different reasons, finding these would exceed the limit of this paper. The maximum value of estimated conditional standard deviation (sigma) is `r round(max(k.ar1garcht21@fit$sigma) * sqrt(252)*100, digits=2)`% and Long-term unconditional level is `r round(unconditional_sigma1 * sqrt(252)*100, digits=2) `%  annualized.

### Annualized conditional standard deviation AR(1)-Garch-t(2,1)
```{r}


sigma.forecast.longrun2 <- ugarchforecast(k.ar1egarcht21, n.ahead = 252)
unconditional_sigma2 <-
  sqrt(
    k.ar1egarcht21@model$pars["omega", 1]/
      (1 -
         k.ar1egarcht21@model$pars["alpha1", 1] -
         k.ar1egarcht21@model$pars["beta1", 1]))
plot(
  c(as.numeric(k.ar1egarcht21@fit$sigma * sqrt(252)),
    as.numeric(sigma.forecast.longrun2@forecast$sigmaFor * sqrt(252))),
  main = "Annaulized Conditional Standard Deviation GARCH",
  type = "l",
  ylab = "sigma annualized")
abline(h = unconditional_sigma2 * sqrt(252), col = "red")

#omega is negative!

max(k.ar1egarcht21@fit$sigma) * sqrt(252)*100

unconditional_sigma2 * sqrt(252)*100
```
In the plot we see the annualized conditional standard deviation of AR(1)-EGARCH-t(2,1). The conditional sigma cannot be estimated because the omega is negative an thus the square-root of the formular for the unconditional is not applicable. A reason for that could be the AR(1) in the model, hence I will try it without.

AR(0)-EGARCH-t(2,1)
```{r}
spec <- 
  ugarchspec(variance.model = list(model = "eGARCH", garchOrder = c(2, 1)),
             mean.model = list(armaOrder = c(0, 0), include.mean = F),
             distribution.model = "std")


k.ar0egarcht21 <- ugarchfit(spec = spec, 
                             data = quotes$quotes.r)

k.ar0egarcht21

```
### Annualized conditional standard deviation AR(0)-EGarch-t(2,1)
```{r}

sigma.forecast.longrun3 <- ugarchforecast(k.ar0egarcht21, n.ahead = 252)
unconditional_sigma3 <-
  sqrt(
    k.ar0egarcht21@model$pars["omega", 1]/
      (1 -
         k.ar0egarcht21@model$pars["alpha1", 1] -
         k.ar0egarcht21@model$pars["beta1", 1]))
plot(
  c(as.numeric(k.ar0egarcht21@fit$sigma * sqrt(252)),
    as.numeric(sigma.forecast.longrun3@forecast$sigmaFor * sqrt(252))),
  main = "Annaulized Conditional Standard Deviation GARCH",
  type = "l",
  ylab = "sigma annualized")
abline(h = unconditional_sigma3 * sqrt(252), col = "red")

#omega is negative!

max(k.ar0egarcht21@fit$sigma) * sqrt(252)*100

unconditional_sigma3 * sqrt(252)*100
```
Even without AR() the result of unconditional sigma is  not applicable. Thus, I decided to choose a different model for the following estimations. 

### Annualized conditional standard deviation Garch(5,5)
```{r}

sigma.forecast.longrun4 <- ugarchforecast(k.garch55, n.ahead = 252)
unconditional_sigma4 <-
  sqrt(
    k.garch55@model$pars["omega", 1]/
      (1 -
         k.garch55@model$pars["alpha1", 1] -
         k.garch55@model$pars["beta1", 1]))
plot(
  c(as.numeric(k.garch55@fit$sigma * sqrt(252)),
    as.numeric(sigma.forecast.longrun4@forecast$sigmaFor * sqrt(252))),
  main = "Annaulized Conditional Standard Deviation GARCH",
  type = "l",
  ylab = "sigma annualized")
abline(h = unconditional_sigma4 * sqrt(252), col = "red")


max(k.garch55@fit$sigma) * sqrt(252)*100

unconditional_sigma4 * sqrt(252)*100
```
In this case the unconditional sigma is applicable. However, it still does not converge to its long run estimation. The maximum value of estimated conditional standard deviation (sigma) is `r round(max(k.garch55@fit$sigma) * sqrt(252)*100, digits=2)`% and Long-term unconditional level is `r round(unconditional_sigma4 * sqrt(252)*100, digits=2) `%  annualized.


## Estimation Garch-Models with out-sample data

(b) estimates of the Value-at-Risk produced by the two models in the out-of-sample period

### Choice of the out-of-the sample data.
```{r}
k <- sum(x)%%6+1
k
```

The result is k=6 hence, the out-of-sample date starts 2020-06-01. 
```{r}
quotes2 <- quotes["2020-05-30/2021-06-01"]
quotes2$quotes.r[is.na(quotes$quotes.r)] <- 0
```


### AR(1)-Garch-t(2,1)
```{r}
spec <- ugarchspec(# variance equation
  variance.model = list(model = "sGARCH", 
                        garchOrder = c(2, 1)),
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE), 
  distribution.model = "std") 

ar1garcht21 <- ugarchfit(spec = spec, 
                           data = quotes2$quotes.r)

ar1garcht21

```
### Garch(55)
```{r}


spec <- ugarchspec(
  variance.model = list(model = "sGARCH", 
                        garchOrder = c(5, 5)),
  mean.model = list(armaOrder = c(0, 0), 
                    include.mean = TRUE), 
  distribution.model = "norm")


garch55 <- ugarchfit(spec = spec, 
                       data = quotes2$quotes.r)

garch55


```
### Standardization of returns.
```{r}


quotes2$rstd <- (quotes2$quotes.r - mean(quotes2$quotes.r, na.rm = T)) /
  sd(quotes2$quotes.r ,na.rm = T)
```

### 1% empirical quantile
```{r}
q01 <- quantile(quotes2$rstd, 0.01, na.rm = T)
q01

```

### Calculation the value-at-risk (VaR)
```{r}

quotes2$VaR1 <- q01 * ar1garcht21@fit$sigma

quotes2$VaR2 <- q01 * garch55@fit$sigma

```

### Plot of returns vs VaR
```{r}
plot(quotes2$quotes.r, 
     main = "returns vs. VaR (GARCH-t)",
     col = "red", lwd = 1, type = 'l', 
     ylim = c(-0.1, 0.1))
abline(h = 0, lty = 2)
lines(quotes2$VaR1, type = 'l', col = "green")
```
In the plot of returns and VaR of AR(1)-GARCH-t(2,1) has been crossed 2 times. 

```{r}

plot(quotes2$quotes.r, 
     main = "returns vs. VaR (GARCH)",
     col = "red", lwd = 1, type = 'l', 
     ylim = c(-0.1, 0.1))
abline(h = 0, lty = 2)
lines(quotes2$VaR2, type = 'l', col = "green")
```

In the plot of returns and VaR of GARCH(5,5) has been crossed 3 times. 

### In how many days losses were higher the assumed VaR?
```{r}
sum(quotes2$quotes.r < quotes2$VaR1) / length(quotes2$VaR1)
```
The number of days were losses are higher than assumed VaR for (AR(1)-GARCH-t(2,1)) is 0.004.  

```{r}
sum(quotes2$quotes.r < quotes2$VaR2) / length(quotes2$VaR2)
```

The number of days were losses are higher than assumed VaR for GARCH(5,5) is 0.008. 



